---
title: "Coronavirus in Mexico"
author: "Enrique A. L贸pez Magall贸n"
#date: '`r paste("Data as of", Sys.Date()-1)`'
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    #social: ["facebook","twitter", "linkedin"]
    #source_code: embed
    vertical_layout: fill
    
---

```{r setup, include=FALSE}
#------------------ Packages ------------------
library(flexdashboard)
library(metathis)
library(dplyr)
library(plotly)
library(stringr)
#install.packages("devtools")
#devtools::install_github("RamiKrispin/coronavirus", force = TRUE)
library(coronavirus)
data(confirmed)
# View(coronavirus)
# max(coronavirus$date)
meta()%>% meta_description("Coronavirus in Mexico overview") %>% meta_viewport()%>% meta_social(
  description = "COVID 19 data for Mexico with map and reference to other countries",
  twitter_site = "@EnriqueALopezM",
  twitter_card_type = "summary_large_image",
  twitter_creator = "@EnriqueALopezM",
  image = "http://www.enadol.de/images/dash.JPG", og_type = "article", og_author = "Enrique Adolfo L贸pez Magall贸n", og_site_name = "http://www.enadol.de/virus/coviddash.html"
  
)
`%>%` <- magrittr::`%>%`
#------------------ Parameters ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
confirmed_color <- "purple"
active_color <- "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "red"
#------------------ Data ------------------
df <- coronavirus %>%
  # dplyr::filter(date == max(date)) %>%
  dplyr::filter(country == "Mexico") %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(-confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(country == "United Arab Emirates", "UAE", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

df_daily <- coronavirus %>%
  dplyr::filter(country == "Mexico") %>%
  dplyr::group_by(date, type) %>%
  dplyr::summarise(total = sum(cases, na.rm = TRUE)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::arrange(date) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(active = confirmed - death - recovered) %>%
  #dplyr::mutate(active = confirmed - death) %>%
  dplyr::mutate(
    confirmed_cum = cumsum(confirmed),
    death_cum = cumsum(death),
    recovered_cum = cumsum(recovered),
    active_cum = cumsum(active)
  )


df1 <- coronavirus %>% dplyr::filter(date == max(date))

conf_df <- coronavirus %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  mutate(parents = "Confirmed") %>%
  ungroup() 


```

Summary
=======================================================================

Row {data-width=400}
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total cumulative confirmed cases",
  icon = "fas fa-user-md",
  color = confirmed_color
)
```
















### death {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$death, na.rm = TRUE) / sum(df$confirmed), 1),
    "%)",
    sep = ""
  ),
  caption = "Deaths (% of confirmed)",
  icon = "fas fa-heart-broken",
  color = death_color
)
```

### recovered {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$recovered, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$recovered, na.rm = TRUE) / sum(df$confirmed), 1),
    "%)",
    sep = ""
  ),
  caption = "Recovered (% of confirmed)",
  icon = "fas fa-heart-broken",
  color = recovered_color
)
```

### active {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df_daily$active, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df_daily$active, na.rm = TRUE) / sum(df$confirmed), 1),
    "%)",
    sep = ""
  ),
  caption = "Active cases (% of confirmed)",
  icon = "fas fa-heart-broken",
  color = active_color
)
```


Row
-----------------------------------------------------------------------

### **Daily cumulative cases by type** (Mexico only)
    
```{r}
plotly::plot_ly(data = df_daily) %>%
  plotly::add_trace(
    x = ~date,
    # y = ~active_cum,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
    # name = "Active",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
   plotly::add_annotations(
     x = as.Date("2021-02-16"),
     y = df_daily$confirmed_cum[392],
     text = paste("Two million cases /
     One million vaccinations 
                 (Feb. 16, 2021)"),
     xref = "x",
     yref = "y",
     arrowhead = 3,
     arrowhead = 1,
     arrowsize = 1,
     showarrow = TRUE,
     ax = -60,
     ay = 140
   ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-19"),
    y = 3,
    text = paste("First death (March 19)"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -90,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-24"),
    y = 14,
    text = paste(
      "First lockdown (March 24)"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 30,
    ay = -120
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-12-24"),
    y = df_daily$confirmed_cum[338],
    text = paste(
      "First vaccination (Dec. 24)"
    ),
    xref = as.Date("2020-12-24"),
    yref = "y",
    arrowhead = 3,
    arrowhead = 1,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -60,
    ay = 120
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-11-14"),
    y = df_daily$confirmed_cum[298],
    text = paste(
      "One million cases (Nov. 14)"
    ),
    xref = as.Date("2020-11-14"),
    yref = "y",
    arrowhead = 3,
    arrowhead = 1,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -60,
    ay = 100
  ) %>% 

  plotly::layout(
    title = "",
    yaxis = list(title = "Cumulative number of cases"),
    xaxis = list(title = "Date"),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
  )
```

Comparison
=======================================================================


Column {.tabset}
-------------------------------------


### **Daily new confirmed cases**
    
```{r}
daily_confirmed <- coronavirus %>%
  dplyr::filter(type == "confirmed") %>%
  dplyr::filter(date >= "2020-02-29") %>%
  dplyr::mutate(country = country) %>%
  dplyr::group_by(date, country) %>%
  dplyr::summarise(total = sum(cases)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = country, values_from = total)

#----------------------------------------
# Plotting the data

daily_confirmed %>%
  plotly::plot_ly() %>%
  plotly::add_trace(
    x = ~date,
    y = ~Mexico,
    type = "scatter",
    mode = "lines+markers",
    name = "Mexico"
  ) %>%
  # plotly::add_trace(
  #   x = ~date,
  #   y = ~France,
  #   type = "scatter",
  #   mode = "lines+markers",
  #   name = "France"
  # ) %>%
  # plotly::add_trace(
  #   x = ~date,
  #   y = ~Spain,
  #   type = "scatter",
  #   mode = "lines+markers",
  #   name = "Spain"
  # ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Peru,
    type = "scatter",
    mode = "lines+markers",
    name = "Peru"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Germany,
    type = "scatter",
    mode = "lines+markers",
    name = "Germany"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Japan,
    type = "scatter",
    mode = "lines+markers",
    name = "Japan"
  ) %>%
  plotly::layout(
    title = "",
    legend = list(orientation="h"),
    yaxis = list(title = "New confirmed cases"),
    xaxis = list(title = ""),
    # paper_bgcolor = "black",
    # plot_bgcolor = "black",
    # font = list(color = 'white'),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```
 
### **Cases distribution by type**

```{r daily_summary}
df_EU <- coronavirus %>%
  # dplyr::filter(date == max(date)) %>%
  dplyr::filter(country == "Mexico" |
    country == "Germany" |
    country == "Peru" |
    country == "Japan") %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(country == "United Arab Emirates", "UAE", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

plotly::plot_ly(
  data = df_EU,
  x = ~country,
  # y = ~unrecovered,
  y = ~ confirmed,
  # text =  ~ confirmed,
  # textposition = 'auto',
  type = "bar",
  name = "Confirmed",
  marker = list(color = active_color)
) %>%
  plotly::add_trace(
    y = ~recovered,
    # text =  ~ death,
    # textposition = 'auto',
    name = "Recovered",
    marker = list(color = recovered_color)
  )%>%
  plotly::add_trace(
    y = ~death,
    # text =  ~ death,
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  ) %>%
  
  plotly::layout(
  
    barmode = "group",
    yaxis = list(title = "Total cases"),
    xaxis = list(title = ""),
    hovermode = "compare",
    legend = list(orientation="h"),
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```

### **Confirmed cases Heatmap**

```{r}
library(d3heatmap)
datain <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", check.names = FALSE)
colnames(datain)[2] <- "Country"
germexper <- filter(datain, datain$Country=="Germany" | datain$Country=="Mexico" | datain$Country=="Peru" | datain$Country == "Japan")
germexper <- germexper[-c(1,3,4)]
germexper[2:ncol(germexper)] <- lapply(germexper[2:ncol(germexper)], as.numeric)

d3heatmap(germexper[2:ncol(germexper)], colors= "Reds", Colv = FALSE, Rowv = FALSE, dendrogram="none", labRow = c("Germany", "Japan", "Mexico", "Peru"), showlegends=FALSE)
```

### **Deaths Heatmap**

```{r}
library(d3heatmap)
datain <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", check.names = FALSE)
colnames(datain)[2] <- "Country"
germexper <- filter(datain, datain$Country=="Germany" | datain$Country=="Mexico" | datain$Country=="Peru" | datain$Country == "Japan")
germexper <- germexper[-c(1,3,4)]
germexper[2:ncol(germexper)] <- lapply(germexper[2:ncol(germexper)], as.numeric)

d3heatmap(germexper[2:ncol(germexper)], colors= "Reds", Colv = FALSE, Rowv = FALSE, dendrogram="none", labRow = c("Germany", "Japan", "Mexico", "Peru"), showlegends=FALSE)
```


### **Death rates (global) as of `r format(max(coronavirus$date), "%b %d, %Y")` **
```{r}
library(plotly)
drates <- read.csv("https://raw.githubusercontent.com/enadol/covid/master/deathrateswo.csv", check.names = FALSE)
plotly::plot_ly() %>% add_bars(x=drates$Country, y=drates$Rates)



```




### **Global cases Treemap**
```{r}

  
  plot_ly(data = conf_df,
          type= "treemap",
          values = ~total_cases,
          labels= ~ country,
          parents=  ~parents,
          domain = list(column=0),
          name = "Confirmed",
          textinfo="label+value+percent parent")
  
```


Map
=======================================================================

### **World map of cases** (*use + and - icons to zoom in/out*)

```{r}
# map tab added by Art Steinmetz
library(leaflet)
library(leafpop)
library(purrr)
library(stringr)
cv_data_for_plot <- coronavirus %>%
  # dplyr::filter(country == "Belgium") %>%
  dplyr::filter(cases > 0) %>%
  dplyr::group_by(country, province, lat, long, type) %>%
  dplyr::summarise(cases = sum(cases)) %>%
  dplyr::mutate(log_cases = 2 * log(cases)) %>%
  dplyr::ungroup()

cv_data_for_plot.split <- cv_data_for_plot %>% split(cv_data_for_plot$type)

hovertext= sprintf("<b> %s Cases: %i </br>Country: %s</br>Province: %s</b>",  str_to_title( cv_data_for_plot$type), cv_data_for_plot$cases, cv_data_for_plot$country, cv_data_for_plot$province)



pal <- colorFactor(c("orange", "red", "green"), domain = c("confirmed", "death", "recovered"))
map_object <- leaflet() %>% setView(lng=0, lat= 22, zoom=3) %>% addProviderTiles(providers$CartoDB.Voyager, providerTileOptions(minZoom=2, maxZoom=10))
names(cv_data_for_plot.split) %>%
  purrr::walk(function(df) {
    map_object <<- map_object %>%
      addCircleMarkers(
        data = cv_data_for_plot,
        lng = ~long, lat = ~lat,
        #                 label=~as.character(cases),
        color = ~ pal(type),
        stroke = FALSE,
        fillOpacity = 0.8,
        radius = ~log_cases/5,
        popup = hovertext,
        #leafpop::popupTable(cv_data_for_plot.split[[df]],
          #feature.id = FALSE,
          #row.numbers = FALSE,
          #zcol = c("type", "cases", "country", "province"),
        group = cv_data_for_plot$type,
        #                 clusterOptions = markerClusterOptions(removeOutsideVisibleBounds = F),
        labelOptions = labelOptions(
          noHide = F,
          direction = "auto"
        )
      )
  })

map_object %>%
  addLayersControl(
    overlayGroups = names(cv_data_for_plot.split),
    options = layersControlOptions(collapsed = FALSE)
  )
```





About
=======================================================================

**The Coronavirus Dashboard**

This Coronavirus dashboard provides an overview of the 2020-2021 Novel Coronavirus COVID-19 (2019-nCoV) epidemic for Mexico. This dashboard is built with R using the R Markdown framework and is based upon this  [dashboard](https://ramikrispin.github.io/coronavirus_dashboard/){target="_blank"} by Rami Krispin.


**Data**

The input data for this dashboard is the dataset available from the [`{coronavirus}`](https://github.com/RamiKrispin/coronavirus){target="_blank"} R package. Make sure to download the development version of the package to have the latest data:

```
install.packages("devtools")
devtools::install_github("RamiKrispin/coronavirus")
```

The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv){target="_blank"}.


**Update**

The data is as of `r format(max(coronavirus$date), "%A %B %d, %Y")` and the dashboard has been updated on `r format(Sys.time(), "%A %B %d, %Y")`.

